Finiteness in Cubical Type Theory

### 2020-07-07 sections/topos.tex
small intro to topos

\chapter{Topos} \label{topos}
In this section we will examine the categorical interpretation of finite sets.
In particular, we will prove that decidable Kuratowski finite types form a
\(\Pi\)-pretopos.
A lot of the work for this proof has been done already: in
Theorem~\ref{cardinal-kuratowski} we saw that Kuratowski finite types were
equivalent to Cardinally finite types.
We will use the latter definition implementation-wise from now on, as it is
slightly easier to work with: CuTT's transport means we can do this without loss
of generality.

\section{Categories in HoTT}
\section{Closure}
\todo{This is just the closure bits pulled over from the finiteness predicates
  section. They need to be reworked to fit in here.}
\subsection{Split Enumerability Closure}
Now that we have a suitable definition of finiteness, we will next prove that
some things are finite.
With the most basic simple types out of the way, the obvious next choice is the
(non-dependent) sums and products: \(\uplus\) and \(\times\).
Both of these types can be constructed from the \emph{dependent} sum, however,
so that is the type we will prove finite.
From that we can derive a much wider array of finiteness proofs.

>>>>

\chapter{Topos} \label{topos}
In this section we will examine the categorical interpretation of finite sets.
In particular, we will prove that decidable Kuratowski finite types form a
\(\Pi\)-pretopos.
A lot of the work for this proof has been done already: in
Theorem~\ref{cardinal-kuratowski} we saw that discrete Kuratowski finite types
were equivalent to Cardinally finite types.
We will use the latter definition implementation-wise from now on, as it is
slightly easier to work with: CuTT's transport means we can do this without loss
of generality.

There are two reasons we're interested in the categorical and topos-theoretic
interpretation of finite sets: first, it's an important theoretical grounding
for finite sets, which allows us to understand them in the context of other
set-like constructions.
Secondly, and more practically, the language of a topos is (or in our case the
\(\Pi\)-pretopos) is a common standard framework for doing mathematics
generally.
This makes it a good basis for an API for building QuickCheck-like generators,
for example.
\section{Categories in HoTT}
\section{Closure}
\todo{This is just the closure bits pulled over from the finiteness predicates
  section. They need to be reworked to fit in here.}
\subsection{Split Enumerability Closure}
Now that we have a suitable definition of finiteness, we will next prove that
some things are finite.
With the most basic simple types out of the way, the obvious next choice is the
(non-dependent) sums and products: \(\uplus\) and \(\times\).
Both of these types can be constructed from the \emph{dependent} sum, however,
so that is the type we will prove finite.
From that we can derive a much wider array of finiteness proofs.
